---
title: "CSDE 502 Winter 2022, Assignment 10"
author: "[Leo Stewart](mailto:lgs17@uw.edu)"
output: 
    bookdown::html_document2:
        number_sections: true
        self_contained: true
        code_folding: hide
        toc: true
        toc_float:
            collapsed: true
            smooth_scroll: false
    pdf_document:
        number_sections: true
        toc: true
        fig_cap: yes
        keep_tex: yes
urlcolor: blue 
---

<!--these following settings control width and also make a DIV tag for black outlines for figures-->
<style type="text/css">
.main-container {
  max-width: 2400px !important;
  margin-left: auto;
  margin-right: auto;
}
</style>

<style>
.outline {
   background-color: #000000;
   padding:1px;
   display: inline-block;
}
</style>

```{r setup, warning=FALSE, message=FALSE}
library(tidyverse)

## For accessing HMD data
library(keyring)
library(HMDHFDplus)

## For table formatting
library(kableExtra)

## For table/figure captions
library(captioner)
figure_nums <- captioner(prefix = "Figure")
table_nums <- captioner(prefix = "Table")

# chunk options
knitr::opts_chunk$set(warning = FALSE, message = FALSE)

# path to this file name
if (!interactive()) {
    fnamepath <- as.character(sys.call(1))[2]
}

## Set plot theme for ggplot figures
theme_set(theme_classic())
```

# Introduction
Births are an essential component of defining and predicting population dynamics. Sex ratio at birth reflects a population's ability to sustain itself and thus understanding sex ratio at birth (SRB) is crucial for understanding the compositional trends of a population. A "common sense" estimate of SRB would usually assume a 50/50 split between males and females (uniformly distributed, completely due to chance). This paper will investigate this hypothesis.

# Methods
Given the large-scale quantitative nature of demographic analysis, identifying trends in sex ratio at birth should be done with statistical analysis. This paper will lay a foundation for such analysis by visualizing trends in SRB for countries in the Human Mortality Database (HMD) data set.

## Data
Data were obtained from the Human Mortality Database (HMD) using the HMDHFDplus R package.

```{r, warning=FALSE, message=FALSE}
## Pull HMD credentials from keychain
HMD_user <- key_list("human-mortality-database")$username
HMD_pass <- key_get(service = "human-mortality-database", 
                    username = HMD_user)

## Declare a function to retrieve data from HMD given a specified data item 
## to pull, convert the dataframe to a tibble, and add a column specifying the 
## item.
get_hmd_data <- function(country, item) {
  readHMDweb(CNTRY = country, 
             item = item,
             username = HMD_user, 
             password = HMD_pass) %>%
    as_tibble() %>%
    add_column(Item = item) %>%
    add_column(Country = country)
}

## Get a list of countries included in the HMD
HMD_countries <- getHMDcountries()

## Get birth data (makes ~50 requests to HMD - takes a bit)
births <- lapply(HMD_countries,
                 function(country) {
                   get_hmd_data(country, "Births")
                 }) %>%
  bind_rows()

## Calculate SRB
births <- births %>%
  mutate(SRB = Male / Female)
```

# Results
This section will present the results of visual analysis.

```{r, warning=FALSE, message=FALSE}
## Plot birth data
ggplot(data = births) +
  geom_line(mapping = aes(x = Year, 
                          y = Total,
                          color = Country)) +
  scale_x_continuous(breaks = ~ seq(floor(.x[1]/25)*25, .x[2], 25)) +
  scale_y_log10(name = "Births, log-10 scale") +
  guides(color = "none") +
  geom_text(data = births %>%
              group_by(Country) %>%
              slice_max(order_by = Year),
            mapping = aes(x = Year, 
                          y = Total,
                          label = Country),
            size = 2,
            nudge_x = 6,
            alpha = 0.75)
```

_`r figure_nums(name = "fig1", caption = "Total births over time (log-10 scale)")`_


`r figure_nums(name = "fig2", display = "cite")` shows a plot of total births over time. As one would expect, some countries have a lot of births and other countries do not. It would be interesting to look at whether there is a relationship between birth count and SRB.

```{r, warning=FALSE, message=FALSE}
## Plot SRB data
ggplot(data = births) +
  geom_line(mapping = aes(x = Year, 
                          y = SRB,
                          color = Country)) +
  scale_x_continuous(breaks = ~ seq(floor(.x[1]/25)*25, .x[2], 25)) +
  scale_y_continuous(name = "Sex Ratio at Birth") +
  guides(color = "none") +
  geom_text(data = births %>%
              group_by(Country) %>%
              slice_max(order_by = Year),
            mapping = aes(x = Year, 
                          y = SRB,
                          label = Country),
            size = 2,
            nudge_x = 6,
            alpha = 0.75) +
  geom_hline(yintercept = 1, lty = 3)
```

_`r figure_nums(name = "fig2", caption = "Sex ratio at birth over time")`_


`r figure_nums(name = "fig2", display = "cite")` shows a plot of SRB for each country over time. It is interesting to see that the average SRB looks closer to 1.05 rather than 1.00. This can be confirmed by calculating the mean of the SRB distribution, where the observed mean is `r round(mean(births$SRB, na.rm = T), digits = 3)`. Some countries vary a lot while others do not.

```{r}
## Plot SRB data for top-3 countries
ggplot(data = subset(births, Country %in% c("USA", "RUS", "JPN"))) +
  geom_point(mapping = aes(x = Total, 
                          y = SRB,
                          color = Country),
             size = 1) +
  scale_x_continuous(name = "Births") +
  scale_y_continuous(name = "Sex Ratio at Birth")
```

_`r figure_nums(name = "fig3", caption = "Sex ratio at birth by birth count for USA, Russia, and Japan")`_


Finally, `r figure_nums(name = "fig3", display = "cite")` shows SRB over time for the three countries with the most births in the most recent year stored in the HMD: USA, Russia, and Japan. Each point represents the SRB for a given year. It looks like SRB varies less for the US and Russia than for Japan.

# Discussion
Visual analysis suggests some relationship between number of births and variance in SRB values. This makes sense, given that countries with a high number of births would likely adhere more closely to an underlying distribution.

# Conclusion
This paper performed exploratory visual analysis on birth data from the Human Mortality Database. Future inquiry could use statistical analysis to further examine the trends surfaced with respect to sex ratio at birth.

<hr>

# {-}

```{r comment=''}
cat(readLines(fnamepath), sep = '\n')
```
