---
title: "CSDE 502 Winter 2022, Assignment 8"
author: "Leo Stewart (netID = lgs17)"
date: '`r format(Sys.time()," %B %d, %Y")`'
email: "lgs17@uw.edu"
output: 
    bookdown::html_document2:
        number_sections: true
        self_contained: true
        code_folding: hide
        toc: true
        toc_float:
            collapsed: true
            smooth_scroll: false
    pdf_document:
        number_sections: true
        toc: true
        fig_cap: yes
        keep_tex: yes
urlcolor: blue 
---

```{r}
library(captioner)
library(tibble)
library(dplyr)
library(stringr)

figure_nums <- captioner(prefix = "Figure")
table_nums <- captioner(prefix = "Table")
```

***Explanation***:
This assignment is intended to give you more practice in manipulating variables. It is a doozy, so complete what you can in a reasonable amount of time!

***Instructions***: 

1. Make sure your Rmd file has no local file system dependencies (i.e., anyone should be able to recreate the output HTML using only the Rmd source file).
1. Make a copy of this Rmd file and add answers below each question. The code that generated the answers should be included, as well as the complete source code for the document.
1. Change the YAML header above to identify yourself and include contact information.
1. For any tables or figures, include captions and cross-references and any other document automation methods as necessary.
1. Make sure your output HTML file looks appealing to the reader.
1. Upload the final Rmd to your github repository.
1. Download [`assn_08_id.txt`](http://staff.washington.edu/phurvitz/csde502_winter_2021/assignments/assn_08_id.txt) and include the URL to your Rmd file on github.com.
1. Create a zip file from your copy of `assn_08_id.txt` and upload the zip file to the Canvas site for Assignment 8. ***The zip file should contain only the text file. Do not include any additional files in the zip file--everything should be able to run from the file you uploaded to github.com. Use zip format and not 7z or any other compression/archive format.***

#
Imagine a new variable: `multirace`, using the following value definitions:

	1 = one race, White
	2 = one race, not White
	3 = two races, includes White
	4 = two races, both non-White
	5 = three or more races, includes White
	6 = three or more races, all non-White
	9 = any race missing (White, Black/African American, American Indian, Asian, other)

## 
**Fill in the codes for the hypothetical cases below (`r table_nums(name = "t01", display = "cite")`).** Assume, as with the other Add Health data, that 6 = "refused" and 8 = "don't know".

`r table_nums(name = "t01", caption = "A hypothetical data set")`

| white| black| AI| asian| raceother|multirace |
|-----:|-----:|--:|-----:|---------:|---------:|
|     1|     0|  0|     0|         0|        1 |
|     0|     1|  0|     0|         0|        2 |
|     1|     0|  0|     1|         0|        3 |
|     0|     1|  1|     0|         0|        4 |
|     1|     1|  0|     1|         0|        5 |
|     0|     1|  0|     0|         1|        4 |
|     0|     1|  1|     0|         1|        6 |
|     1|     0|  1|     0|         0|        3 |
|     1|     1|  1|     0|         1|        5 |
|     6|     1|  8|     1|         6|       NA |	

## 
**Using this data frame (code below), report how many cases checked more than one race.** Use R code to make this calculation and use inline expressions.

```{r}
dat <- 
structure(
    list(
        white = c(1L, 0L, 1L, 0L, 1L, 0L, 0L, 1L, 1L, 6L),
        black = c(0L, 1L, 0L, 1L, 1L, 1L, 1L, 0L, 1L, 1L), 
        AI = c(0L, 0L, 0L, 1L, 0L, 0L, 1L, 1L, 1L, 8L), 
        asian = c(0L, 0L, 1L, 0L, 1L, 0L, 0L, 0L, 0L, 1L), 
        raceother = c(0L, 0L, 0L, 0L, 0L, 1L, 1L, 0L, 1L, 6L), 
        multirace = c(NA, NA, NA, NA, NA, NA, NA, NA, NA, NA)
    ), 
    class = "data.frame", 
    row.names = c(NA, -10L)
)
```

`r sum(rowSums(sign(dat), na.rm = TRUE) > 1)` respondents checked more than one race, although this includes the last case, which looks somewhat suspect. Excluding this case leaves `r sum(rowSums(sign(dat), na.rm = TRUE) > 1)-1` respondents.

##
**Write R code to create the multirace variable, using the data set [AHwave1_v3.rds](http://staff.washington.edu/phurvitz/csde502_winter_2021/data/AHwave1_v3.rds).** _Hint: You may want to create another variable, numrace, that counts the number of races._ Use [`download_file()`](https://stat.ethz.ch/R-manual/R-devel/library/utils/html/download.file.html) and `tempdir()` to download the file to a temporary directory to avoid local file system dependencies.

```{r}
## URL to RDS file
rds_link <- str_c("http://staff.washington.edu/phurvitz/",
                  "csde502_winter_2021/data/AHwave1_v3.rds")
## Locate a path to a temporary directory
td <- tempdir()
## Create a new file path to download the RDS file to
rds_file <- file.path(td, "AHwave1_v3.rds")
## Download the file to the new file path
curl::curl_download(url = rds_link, destfile = rds_file)
## Read RDS into environment
ahw1_dta <- readRDS(rds_file)

## Variables to pull from DTA
dtavars <- c("white", "black", "AI", "asian", "raceother")

## Function to iterate through the variable names and add them to the dataframe
dta_to_df <- function(dtavars, sanitize_labels) {
  
  dt <- lapply(dtavars, function(v) {
    ## Pull out the variable values
    dtcol <- ahw1_dta[[v]]
    ## Pull out the variable metadata
    meta <- attributes(dtcol)
    ## Convert vector and metadata object to a factor in a tibble,
    ## format the labels nicely, and add a description to the column 
    ## as a label
    tibble(factor(dtcol, 
                  labels = sanitize_labels(names(meta$labels)), 
                  ordered = TRUE)) %>%
      purrr::set_names(nm = v) %>%
      { attr(.[[v]], "label") <- meta$label; . }
  }) %>%
    ## Bind all the individual columns together
    bind_cols()
}

## Use the function to pull specified variables from AHwave1 DTA
ahw1a <- dta_to_df(dtavars, function(lbls) {
  lbls %>% 
    str_remove(regex("\\(.*\\)\\ ")) %>% 
    str_remove(" \\(.*")
})

## Calculate multirace column
ahw1a <- ahw1a %>%
  ## Decide how much each response category will count to the total sum
  ## Currently, Refused and Don't know are not counted towards the sum (i.e.
  ## counted as zeros)
  mutate(across(everything(), ~ case_when(.x == "Not marked" ~ 0,
                                          .x == "Marked" ~ 1,
                                          .x == "Refused" ~ 0,
                                          .x == "Don't know" ~ 0))) %>%
  ## Add a row to connect each observation to its respondent
  add_column(obsid = 1:nrow(ahw1a)) %>%
  ## Reshape long to make summing easier
  tidyr::pivot_longer(cols = -obsid) %>%
  ## Group by respondent and sum to count how many races were marked
  ## by each respondent
  group_by(obsid) %>%
  add_count(obsid, wt = value, name = "numrace") %>%
  ungroup() %>%
  ## Return data to wide format
  tidyr::pivot_wider(id_cols = c(obsid, numrace), 
                     names_from = "name", 
                     values_from = "value") %>%
  ## Create multirace variable based on the column we just created
  ## and the responses to individual categories
  mutate(multirace = case_when(numrace == 1 & white == 1 ~ 1,
                               numrace == 1 & white == 0 ~ 2,
                               numrace == 2 & white == 1 ~ 3,
                               numrace == 2 & white == 0 ~ 4,
                               numrace >= 3 & white == 1 ~ 5,
                               numrace >= 3 & white == 0 ~ 6,
                               numrace == 0 ~ 9)) %>%
  ## Remove the intermediate columns
  select(-obsid, -numrace)
```

## 
**Label the `multirace` variable as well as its values using attribute labels.**.

```{r}
## Convert numbers to a factor with labels
ahw1a$multirace <- factor(ahw1a$multirace,
                         levels = c(1:6, 9),
                         labels = c("one race, white", 
                                    "one race, not white", 
                                    "two races, including white", 
                                    "two races, none white", 
                                    "three or more races, including white", 
                                    "three or more races, none white", 
                                    "race missing"))

## Add column-level label
attr(ahw1a$multirace, "label") <- str_c("Synthetic column representing ",
                                       "multiracial background of respondents")
```

## 
**Include below a contingency table of the multirace variable. Make sure that the values are labelled so the table is readable, and also include any missing values.**

```{r}
## Print out contingency table (honestly not sure what a contingency table is
## in this context, but hope this works)
table(ahw1a$multirace) %>%
  kableExtra::kable(col.names = c("multirace", "Count"), format = "html") %>%
  kableExtra::kable_styling()
```

# 
**Review part B of each of the answers (i.e., `H1KQ1B .. H1KQ10B`) to the Knowledge Quiz (Section 19 of the Add Health questionnaire, documented in `INH19PUB.PDF`). The 10 questions each ask: "How confident are you that your answer is correct?"**

## 
**Write R code that creates a single summary variable named `kqconfidence`, with a larger number representing the respondent being more confident across all questions (scale of 0 to 3 for each individual question; `kqconfidence` will be the sum for each subject across the 10 questions). Note that any observations with value 7 (i.e., age less than 15) should be removed from the data frame, and values 6, 8, and 9 should be coded as `NA` (i.e., missing) for the purposes of scoring confidence. Document your code so that the reader knows how you scored the scale and how you handled missing values. Make sure to label the new variable.**

```{r}
## Variables to pull from DTA
dtavars <- str_c("h1kq", 1:10, "b")

## Create table with knowledge quiz variables
ahw1b <- dta_to_df(dtavars, function(lbls) {
  lbls %>% 
    str_remove(regex("\\(.*\\)\\ ")) %>% 
    str_remove(" \\(.*")
}) %>%
  ## Add column to keep track of which response belongs to which response
  ## before reshaping data
  add_column(obsid = 1:nrow(.)) %>%
  ## Reshape longer
  tidyr::pivot_longer(cols = -obsid) %>%
  ## Remove all value 7 responses
  filter(value != "Legitimate skip") %>%
  ## Convert values to a number as follows:
  ## 1 (Very)             --> 3
  ## 2 (Moderately)       --> 2
  ## 3 (Slightly)         --> 1
  ## 4 (Not at all)       --> 0
  ## 5 (Refused)          --> -1
  ## 7 (Don't know)       --> -3
  ## 8 (Not applicable)   --> -4
  ## This is done somewhat hackily by subtracting 4 from all values
  mutate(value = 4 - as.numeric(value)) %>%
  ## After converting, keep non-negative values (Not at all .. Very) and replace
  ## negative values (Refused .. Not applicable) with NA values
  mutate(value = case_when(value >= 0 ~ value,
                           value <= 0 ~ NA_real_)) %>%
  ## Create kqconfidence by summing together individual scores. NA values are
  ## very crudely replaced with zeros in the sum
  group_by(obsid) %>%
  add_count(obsid, wt = tidyr::replace_na(value, 0), name = "kqconfidence") %>%
  ungroup() %>%
  ## Restore to wide format
  tidyr::pivot_wider(id_cols = c(obsid, kqconfidence), 
                     names_from = "name", 
                     values_from = "value") %>%
  ## Remove obsid column since we don't need it any more
  select(-obsid)

attr(ahw1b$kqconfidence, "label") <- str_c("Sum of knowledge quiz confidence",
                                           "variables, where a higher score ",
                                           "represents a higher score.")
```

## 
**Create and include below a contingency table from `kqconfidence` with raw counts, percentages, and cumulative percentages.**
```{r}
## Creating contingency table
ahw1b %>%
  ## Group by each possible kqconfidence score
  group_by(kqconfidence) %>%
  ## For each score, count how many people got that score
  summarise(Count = n()) %>%
  ## Convert from a summary to a new table
  as_tibble() %>%
  ## Convert counts to percentage of total respondents
  mutate(Percentage = Count*100 / sum(.$Count)) %>%
  ## Create cumulative sum of percentage column
  mutate(`Cumulative Percentage` = cumsum(Percentage)) %>%
  ## Round everything to 2 significant digits
  mutate(across(everything(), ~ signif(.x, digits = 2))) %>%
  ## Format the table and print it out
  kableExtra::kable("html") %>%
  kableExtra::kable_styling()
```

## 
**[BONUS] For each subject there were zero to 10 _missing_ answers to each of the 10 component questions. We would like to know what this distribution is. Include below a table that shows the count of subjects for each unique value of the count of missing questions.**

## 
**For each possible value of the Knowledge Quiz Part A sum score (from [Lesson 8](http://staff.washington.edu/phurvitz/csde502_winter_2021/week8.html#scoring-across-columns)) what is the mean `kqconfidence` level?**

## 
**[BONUS] For each respondent, create two different confidence scores: a confidence score for the items answered correctly and a confidence score for the items answered incorrectly. How many respondents are more confident when answering incorrectly?**

# Source code
```{r, comment='', echo=FALSE}
fname <- knitr::current_input()
cat(readLines(fname), sep = "\n")
```
