---
title: "Assignment 7"
author: 'Leo Stewart (lgs17@uw.edu)'
date: '`r format(Sys.time()," %B %d, %Y")`'
email: "lgs17@uw.edu"
output:
  bookdown::html_document2:
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
    toc_depth: 6
    number_sections: true
    self_contained: yes
---


```{r setup, message=FALSE}
rm(list = ls())
library(tidyverse)

## For downloading and unzipping .dta file
library(curl)
library(utils)

## For table and figure captions
library(captioner)
table_nums <- captioner(prefix = "Table")
figure_nums <- captioner(prefix = "Figure")

## For reading in .dta file
library(haven)

## Bug in current kableExtra package where collapse_rows doesn't work,
## so using dev build
devtools::install_github(repo="haozhu233/kableExtra", ref="a6af5c0", 
                         force = FALSE, quiet = TRUE)
## For table formatting
library(kableExtra)
```

# Generating A New Dataframe

## Downloading and unzipping .dta file
```{r}
## Path to zip file
zip_link <- str_c("http://staff.washington.edu/phurvitz/",
                  "csde502_winter_2021/data/21600-0001-Data.dta.zip")

## Locate a path to a temporary directory
td <- tempdir()
## Create a new file path to download the zip file to
zip_file <- file.path(td, "data.dta.zip")
## Download the file to the new file path
curl_download(url = zip_link, destfile = zip_file)

## Create a directory path to extract the files to
extracted_dir <- file.path(td, "extracted")
## Initialize the directory if it doesn't exist already
if (!dir.exists(extracted_dir)) dir.create(extracted_dir)
## Extract the contents into the directory
unzip(zipfile = zip_file, exdir = extracted_dir)

## Locate the downloaded and extracted .dta file
dta_path <- file.path(extracted_dir, "21600-0001-Data.dta")
```

## Loading .dta file into R
```{r}
## Read in the .dta file
AHwave1 <- read_dta(dta_path)

## Names of variables to pull
dtavars <- c("BIO_SEX", str_c("H1NB", 1:7))

## Function to remove level numbers in parenthesis from factor labels
sanitize_labels <- function(lbls) str_remove(lbls, regex("\\(.\\)\\ "))

## Iterate through the variable names and add them to the dataframe
ahdt <- lapply(dtavars, function(v) {
  ## Pull out the variable values
  dtcol <- AHwave1[[v]]
  ## Pull out the variable metadata
  meta <- attributes(dtcol)
  ## Convert vector and metadata object to a factor in a tibble,
  ## format the labels nicely, and add a description to the column 
  ## as a label
  tibble(factor(dtcol, 
                labels = sanitize_labels(names(meta$labels)), 
                ordered = TRUE)) %>%
    set_names(nm = v) %>%
    { attr(.[[v]], "label") <- meta$label; . }
}) %>%
  ## Bind all the individual columns together
  bind_cols() %>%
  ## Add an overall description to the table as a label
  { attr(., "label") <- str_c("Sex category and neighborhood data ",
                              "from Add Health Wave 1, version 1"); . }
```


# Saving RDS Output
```{r}
## Declare a file path in the original temporary directory to save 
## the .RDS file in
RDSpath <- file.path(td, "ahdata.RDS")
## Save the .RDS file at the file path
saveRDS(ahdt, RDSpath)
```


# Frequency Tables
## Single variable tables
```{r}
## Create a function that groups the Add Health data subset by
## the given set of variables, calculates a frequency table, and
## renames the grouping variable(s) column(s) to a generic name
## so that they can be consistently bound together
freq_table <- function(varnames) {
  ## Create frequency table:
  ft <- ahdt %>%
    ## Group by given variables and add count and percent columns
    with_groups(all_of(varnames), 
                summarise, 
                Count = n(),
                Percent = str_c(round(n()*100 / nrow(ahdt), 
                                      digits = 1), "%")) %>%
    ## Add a column recording what the grouping variables were
    add_column(Group = str_c(varnames, collapse = ", ")) %>%
    ## Convert from summary table to tibble
    as_tibble()
  
  ## Rename specified varnames to Var1, Var2, ...
  varidx <- which(names(ft) %in% varnames)
  names(ft)[varidx] <- str_c("Var", varidx)
  
  ## Convert response variable column(s) from factors to characters
  ft %>%
    mutate(across(starts_with("Var"), as.character))
}

## Make a combined frequency table:
dtavars %>%
  ## Make a frequency table for each individual variable
  lapply(freq_table) %>%
  ## Bind them all together
  bind_rows() %>%
  ## Select the columns for the output table
  select(Group, Response = Var1, Count, Percent) %>%
  ## Style and print table
  kable("html") %>%
  kable_styling() %>%
  collapse_rows(columns = 1)
```
`r table_nums(name = "ft1", caption = "Combined frequency table for sex category and neighborhood variables")`

## Two-variable frequency table
```{r}
## Make a frequency table for each combination of H1NB1 and H1NB5
freq_table(c("H1NB1", "H1NB5")) %>%
  ## Give descriptive names
  select(`You know most of the people in your neighborhood` = Var1,
         `Do you usually feel safe in your neighborhood?` = Var2, 
         Count, 
         Percent) %>%
  ## Style and print table
  kable("html") %>%
  kable_styling()
```
`r table_nums(name = "ft2", caption = "Frequency table for variables H1NB1 and H1NB5")`

  
4.Creating Graph From Dataframe
```{r}
## Create plot
ggplot(ahdt, mapping = aes(x = H1NB1, 
                           fill = BIO_SEX)) +
  ## Add side-by-side bars (instead of stacked bars)
  geom_bar(position = "dodge") +
  ## Use classic visual theme
  theme_classic() +
  ## Add a chart title
  ggtitle("You know most of the people in your neighborhood")
```
`r figure_nums(name = "fig1", caption = "Responses to H1NB1 separated by sex category")`

# Source code
```{r, comment='', echo=FALSE}
fname <- knitr::current_input()
cat(readLines(fname), sep = "\n")
```