---
title: "Assignment 5"
author: 'Leo Stewart (lgs17@uw.edu)'
date: '`r format(Sys.time()," %B %d, %Y")`'
email: "lgs17@uw.edu"
output:
  bookdown::html_document2:
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
    toc_depth: 6
    number_sections: true
    self_contained: yes
---


```{r setup, message=FALSE}
library(tidyverse)

## for accessing HMD in Q1
library(keyring)
library(HMDHFDplus)

## for table formatting
library(kableExtra)

## For table and figure captions
library(captioner)
table_nums <- captioner(prefix = "Table")
figure_nums <- captioner(prefix = "Figure")
```

<!-- Create a Rmd file in your local file system that generates HTML with at least two interesting summaries (tabular and/or graphical) of one or more time series data sets from the Human Fertility Database. -->



```{r, fig.align='left'}
## Pull HMD credentials from keychain
HMD_user <- key_list("human-mortality-database")$username
HMD_pass <- key_get(service = "human-mortality-database", 
                    username = HMD_user)

## Declare a function to retrieve data from HMD given a specified data item 
## to pull, convert the dataframe to a tibble, and add a column specifying the 
## item.
get_aut_data <- function(item) {
  readHMDweb(CNTRY = "AUT", 
             item = item,
             username = HMD_user, 
             password = HMD_pass) %>%
    as_tibble() %>%
    add_column(Item = item)
}

## Get birth data 
aut_births <- get_aut_data(item = "Births")

## Plot birth data
ggplot(data = aut_births) +
  geom_line(mapping = aes(x = Year, y = Total)) +
  theme_classic() +
  scale_x_continuous(breaks = ~ seq(floor(.x[1]/10)*10, .x[2], 10))
```
_`r figure_nums(name = "fig_births", caption = "Births, 1871-2019, Austria")`_

```{r}
aut_births %>%
  mutate(Decade = floor(Year / 10)*10) %>%
  mutate(Decade = str_c(Decade, "-", Decade+9)) %>%
  group_by(Decade) %>%
  summarise(Total = formatC(sum(Total), big.mark=",")) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", 
                                      "responsive"), 
                full_width = FALSE)
```
_`r table_nums(name = "tab_births", caption = "Births by decade, 1871-2019, Austria")`_

`r figure_nums(name = "fig_births", display = "cite")` shows a very simple plot; however the massive changes in births over time are interesting nonetheless. Besides obvious shocks, like the WWI and WWII, I would guess that some of the change is attributable to changes in political boundaries as a result of WWI. I do not know much about European history beyond this, but I am curious as to the jaggedness in the 1940s and the smaller peaks in the 1980s and 90s. The same trend can be interpreted in tabular format, as shown in `r table_nums(name = "tab_births", display = "cite")`.

